---
title: "HW 2"
author: "Eve Wicksteed"
date: "22/09/2019"
output: 
  html_document:
    keep_md: true
    toc: false
    number_sections: false
    toc_float: true
---

```{r load, warning=FALSE, echo=FALSE}
library(gapminder)
library(tidyverse)
library(ggplot2)
```

## Intro

Have a look at the gapminder data:
```{r}
DT::datatable(
gapminder)
```

## Question 1.1

Use filter() to subset the gapminder data to three countries of your choice in the 1970’s.

```{r q1.1}
knitr::kable(
(eves_data <- gapminder %>%
  filter((country == "South Africa" | country == "Canada" | country == "United Kingdom") & 
           (year >= 1970 & year <= 1979)))
)

```


## Question 1.2

Use the pipe operator %>% to select “country” and “gdpPercap” from your filtered dataset in 1.1.

```{r q1.2}
knitr::kable(
eves_data %>% 
  select(country, gdpPercap))
  
```

## Question 1.3

Filter gapminder to all entries that have experienced a drop in life expectancy. 
Be sure to include a new variable that’s the increase in life expectancy in your 
tibble. Hint: you might find the lag() or diff() functions useful.

First attempt before class on Tuesday... see later answer. 
```{r q1.3}

knitr::kable(
gapminder %>%
  mutate(change_lifeExp = lifeExp - first(lifeExp)) %>% 
  filter(change_lifeExp <0)
)

```

After class on Tuesday (24 Sep):
```{r}
gapminder %>% 
  arrange(year) %>% 
  group_by(country) %>%
  mutate(gdpPercap_inc = gdpPercap - lag(gdpPercap)) %>% 
  filter(gdpPercap_inc < 0)

```






## Question 1.4

Filter gapminder so that it shows the max GDP per capita experienced by 
each country. Hint: you might find the max() function useful here.

```{r q1.4}

DT::datatable(
gapminder %>%
  group_by(country) %>% 
  filter(gdpPercap == max(gdpPercap)))


```


## Question 1.5

Produce a scatterplot of Canada’s life expectancy vs. GDP per capita using ggplot2, 
without defining a new variable. That is, after filtering the gapminder data set, 
pipe it directly into the ggplot() function. Ensure GDP per capita is on a log scale.

```{r q1.5, fig.width = 5, fig.height = 2}

gapminder %>% 
  filter(country =="Canada") %>%
  ggplot(aes(gdpPercap, lifeExp)) +
    geom_point() +
    xlab("GDP per capita") +
    ylab("Life expectancy") +
    ggtitle("Life expectancy vs. GDP per capita in Canada")+
    scale_x_log10()
  

```



##Question 2
Pick one categorical variable and one quantitative variable to explore. Answer 
the following questions in whichever way you think is appropriate, using dplyr:

What are possible values (or range, whichever is appropriate) of each variable?
What values are typical? What’s the spread? What’s the distribution? Etc., tailored to the variable at hand.
Feel free to use summary stats, tables, figures.

### A look at life expectancy:

Summary stats:

The range of values are from 23.60 to 82.60. 
```{r q2}
summary(gapminder$lifeExp)

```

Distribution of life expectancy
```{r}
ggplot(gapminder, aes(lifeExp)) +
  geom_histogram()+
  ylab("Count")+
  xlab("Life expectancy")+
  ggtitle("Distribution of life expectancy for all data")
```




### A look at continents:

These are possible values and the count of each:
```{r}
summary(gapminder$continent)
```

And the histogram for continents:
```{r}
ggplot(gapminder, aes(continent)) +
  geom_bar()+
  ylab("Count")+
  xlab("Continent")
```


Then we can also look at the distribution for life expectancy over the continents:
```{r}
ggplot(gapminder, aes(continent, lifeExp)) +
  geom_point() +
  geom_boxplot()
```


## Question 3

Make two plots that have some value to them. That is, plots that someone might actually consider making for an analysis. Just don’t make the same plots we made in class – feel free to use a data set from the datasets R package if you wish.

A scatterplot of two quantitative variables.
One other plot besides a scatterplot.
You don’t have to use all the data in every plot! It’s fine to filter down to one country or a small handful of countries.

Load in the air quality dataset from r datasets which shows New York air quality measurements
```{r}
library(datasets)

DT::datatable(airquality)
```

Plot of temperature vs. ozone

```{r}

airquality %>% 
  ggplot(aes(Temp, Ozone)) +
    geom_point(aes(size = Wind), alpha = 0.3) +
    xlab("Temperature") +
    ylab("Ozone concentration") +
    ggtitle("Temperature vs. Ozone concentration in New York")
```


Another plot:

```{r}


airquality %>% 
  ggplot(aes(Ozone)) +
  geom_density(aes(fill=factor(Month)),alpha=0.8)+ 
  labs(title="Density plot", 
       subtitle="Ozone concentrations by month",
       caption="Source: airquality",
       x="Ozone concentration",
       fill="Month")

```


## Recycling

For people who want to take things further.

Evaluate this code and describe the result. Presumably the analyst’s intent was to get the data for Rwanda and Afghanistan. Did they succeed? Why or why not? If not, what is the correct way to do this?


```{r}
DT::datatable(filter(gapminder, country == c("Rwanda", "Afghanistan")))
```

This code does work and produces a table with the data for Rwanda and Afghanistan.Presumably this is because `country == c("Rwanda", "Afghanistan")` evaluated the same as `(country == "Rwanda" | country=="Afghanistan")`, thus data for both countries is returned. 


