---
title: "HW 05"
author: "Eve Wicksteed"
date: "12/10/2019"
output: 
  html_document:
    keep_md: true
    theme: paper
---

```{r, echo = FALSE, warning = FALSE, message = FALSE}
library(tidyverse)
library(gapminder)
library(dplyr)
library(DT)
library(knitr)
knitr::opts_chunk$set(fig.width=4, fig.height=3, warning = FALSE, fig.align = "center")
```

# Exercise 1: Explain the value of the here::here package (10%)


The here::here package is useful for multiple reasons. Often people have different operating systems on their computers which can cause problems when setting file paths. For example, Windows uses a backslash ( \ ) where MacOS uses a forwardslash (/). 
For example, my current working directory is: "Users/Eve/courses/STAT_545/STAT545-participation" but this wouldn't be the same on a Windows computer. 

Here::here is also useful because it makes it easier to use subdirectories. Once can set the subdirectory to run code from without being in that directory. 


#Exercise 2: Factor management (20%)
Task: Choose one dataset (of your choice) and a variable to explore. After ensuring the variable(s) you’re exploring are indeed factors, you should:

First print the gapminder dataset to see which columns are factors:
```{r, warning = FALSE, message = FALSE}
gapminder
```
From this I can see that "country" and "continent" are factors. 

From this I'm going to choose 5 countries as factors and drop the other ones that I'm not using
```{r, warning = FALSE, message = FALSE}
(gcountries_dropped <- gapminder %>% 
  filter(country == c("South Africa", "Netherlands", "Canada", "United States", "Zambia")) %>% # choose only certain factors
  droplevels()) # drop the unused factors
```

